---
title: "R Notebook"
output: html_notebook
---

```{r message=F}
library(readr)
library(tidyr)
library(stringr)
library(vroom)
library(dplyr)
library(networkD3)

source("app3_sankey_app/functions_literature.R")

data_location = '../results/literature_outputs/'
```

# find molecular traits that can be looked at 
```{r}
lit_counts <- read_tsv("../results/literature_outputs/lit_space_statsV3.tsv")

inputs<- readRDS("../../01_MR_related/scripts/app2_heatmaps_app/data/inputsV3.rds")
mr_data <- inputs$merged %>% select(-c(`BCAC 2020`:`TNBC`), -label, - exposure_name) %>% 
                            mutate(non_zero = rowSums(.[4:6]!=0)) %>% 
                            filter(non_zero > 0) %>% 
                            filter(exposure_cat %in% c("Proteins", "Metabolites", "Lipids"))

mr_lit <- left_join(mr_data, lit_counts %>% select('id.exposure', "lit_space_size_tidy",  "unique_triples"  ,"unique_pairs"  )
                    , by = "id.exposure") %>% 
          filter(!is.na(lit_space_size_tidy)) %>% 
          filter(lit_space_size_tidy > 49) %>% 
          arrange(-non_zero) %>% distinct()

write_tsv(mr_lit, "../results/trait_with_mr_and_lit.tsv")
```


# Breast cancer
```{r message=F}
# tidy breast cancer space and extract sequential triples

## prod data
dat<- vroom("../results/literature_outputs/breast_cancer_litspace_prod.csv",show_col_types = FALSE) %>%  mutate(lit.id = as.character(lit.id))

# tidy space 
bc_triples_tidy_count <- tidy_lit_space(dat) %>%    
    mutate(term1.type_verbose  = ifelse(grepl("obesity", term1, ignore.case = T), 'antro', term1.type_verbose)) %>%
    mutate(term2.type_verbose  = ifelse(grepl("obesity", term2, ignore.case = T), 'antro', term2.type_verbose))

write_csv(bc_triples_tidy_count, "../results/literature_outputs/breast_cancer_litspace_prod_tidy_space.csv")

bc_triples <- get_breast_cancer_triples(bc_triples_tidy_count)
```



# Molecular traits 

For each molecular trait (i.e. trait with anchor):
	 - load tidy lit space form `lit_spaces_finalset_tidy.RData`
	 - extract and link trait triples
	 - make mini sankey of trait triples (for exploratory purposes)
	 - overlap literature spaces of trait and breast cancer
	 - the output has the full_sankey or subset_sankey (excludes n=1 (or more)) 
	 - save overlap intermediate terms


## HDL cholestrol

```{r}

load(paste0(data_location, "lit_spaces_finalset_tidyV3.RData"))
id = 'ukb-d-30760_irnt'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c('High Density Lipoproteins'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)

KEY_TERM= 'High Density Lipoproteins'


out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=2, sankey_font = 13,bc_triples)
out$full_sankey
out$subset_sankey

#write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), "NEW.csv")) # already saved!

```

#### path finder
```{r}
# find path when term is the link between two spaces
mediator = "testosterone"

AB = trait_triples$joined_triples %>% filter(term2 == mediator)
TA = trait_triples$joined_triples %>% filter(term2 %in% AB$term1)
from_trait <- bind_rows(AB, TA) %>% rename(sankey_n=n)

from_BC <- tibble()

links<- mediator
for (i in rev(1:length(bc_triples))){ # start with the inner
  tmp <- bc_triples[[i]] %>% filter(term1 %in%links)
  links <- c(links, tmp$term2) # in the next triple level collect next triple in chain
  from_BC <- bind_rows(from_BC, tmp)
}

left_join(from_trait, trait_tidy) %>% arrange(term1, term2) %>% View()

trait_tidy %>% filter(term1 %in% from_trait$term1, term2 %in% from_trait$term2) %>% arrange(term1, term2) %>% View()

mini <- overlap_trait_and_bc( bind_rows(AB, TA) , KEY_TERM =KEY_TERM, n_filter=0, sankey_font = 13,bc_triples)

```

```{r}
# find path when term is from BC space
mediator = "CATHEPSIN D"

from_BC <- tibble()

links<- mediator

# triple 3:

t3 <- bc_triples$triple3 %>% filter(term2 == mediator)
t2 <- bc_triples$triple2 %>% filter(term1 == mediator)
t1 <- bc_triples$triple1 %>% filter(term1 %in% t2$term2)
t4 <- bc_triples$triple4 %>% filter(term2 %in% t3$term1)

from_BC <- bind_rows(t1,t2,t3,t4)


AB = trait_triples$joined_triples %>% filter(term2 %in% from_BC$term1)
TA = trait_triples$joined_triples %>% filter(term2 %in% AB$term1)
from_trait <- bind_rows(AB, TA) %>% rename(sankey_n=n)


left_join(from_trait, trait_tidy) %>% arrange(term1, term2) %>% View()
left_join(from_BC, bc_triples_tidy_count) %>% arrange(term1, term2) %>% View()

trait_tidy %>% filter(term1 %in% from_trait$term1, term2 %in% from_trait$term2) %>% arrange(term1, term2) %>% View()

mini <- overlap_trait_and_bc( bind_rows(AB, TA) , KEY_TERM =KEY_TERM, n_filter=0, sankey_font = 13,bc_triples)

```
```{r}
# find path when term is from BC space
mediator = "CATHEPSIN D"

from_BC <- tibble()

links<- mediator

# triple 3:

t3 <- bc_triples$triple3 %>% filter(term2 == mediator)
t2 <- bc_triples$triple2 %>% filter(term1 == mediator)
t1 <- bc_triples$triple1 %>% filter(term1 %in% t2$term2)
t4 <- bc_triples$triple4 %>% filter(term2 %in% t3$term1)

from_BC <- bind_rows(t1,t2,t3,t4)


AB = trait_triples$joined_triples %>% filter(term2 %in% from_BC$term1)
TA = trait_triples$joined_triples %>% filter(term2 %in% AB$term1)
from_trait <- bind_rows(AB, TA) %>% rename(sankey_n=n)


left_join(from_trait, trait_tidy) %>% arrange(term1, term2) %>% View()
left_join(from_BC, bc_triples_tidy_count) %>% arrange(term1, term2) %>% View()

trait_tidy %>% filter(term1 %in% from_trait$term1, term2 %in% from_trait$term2) %>% arrange(term1, term2) %>% View()

mini <- overlap_trait_and_bc( bind_rows(AB, TA) , KEY_TERM =KEY_TERM, n_filter=0, sankey_font = 13,bc_triples)

```

## Calpastatin

```{r}

load(paste0(data_location, "lit_spaces_finalset_tidy.RData"))
id = 'prot-a-366'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c('calpastatin'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)


df<- trait_triples$joined_triples %>% 
      mutate(term1 = ifelse(term1 == 'calpain inhibitor', "calpastatin", term1)) %>% 
      filter(term2 != "calpain inhibitor")
KEY_TERM= 'calpastatin'


out <- overlap_trait_and_bc(df, KEY_TERM =KEY_TERM, n_filter=1, sankey_font = 13,bc_triples) # NB using n=1
out$full_sankey
out$subset_sankey

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```

## Haptoglobin


```{r}

load(paste0(data_location, "lit_spaces_finalset_tidyV3.RData"))
id = 'prot-a-1369'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_tidy <- trait_tidy %>% 
  mutate(term1  = ifelse(term1 %in% c("HP", "Haptoglobin 1-1", 
                                     "Haptoglobin 2-1","Haptoglobin 2-2", 
                                     "Haptoglobins"), "Haptoglobin", term1)) %>% 
  mutate(term2  = ifelse(term2 %in% c("HP", "Haptoglobin 1-1", 
                                     "Haptoglobin 2-1","Haptoglobin 2-2", 
                                     "Haptoglobins"), "Haptoglobin", term2)) %>% distinct()
                           

trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c('Haptoglobin'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)

KEY_TERM= 'Haptoglobin'


out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=2, sankey_font = 13,bc_triples)
out$full_sankey
out$subset_sankey

#write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```




## IL3R

```{r}

load(paste0(data_location, "lit_spaces_finalset_tidy.RData"))
id = 'prot-a-1530'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_tidy <- trait_tidy %>% mutate(term1  = ifelse(term1 == "Interleukin-3", "IL3RA", term1)) %>% 
                             mutate(term1  = ifelse(term1 == "Interleukin-3 Receptor", "IL3RA", term1)) %>% 
                             mutate(term2  = ifelse(term2 == "Interleukin-3", "IL3RA", term2)) %>% 
                             mutate(term2  = ifelse(term2 == "Interleukin-3 Receptor", "IL3RA", term2)) 

trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c('IL3RA'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)


KEY_TERM= 'IL3RA'


out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=1, sankey_font = 13,bc_triples) # NB using n=1
out$full_sankey
out$subset_sankey

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```

## FABP4

```{r}

load(paste0(data_location, "lit_spaces_finalset_tidy.RData"))
id = 'prot-b-71'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c('FABP4', "Fatty Acid-Binding Proteins", "Fatty Acid-Binding Protein 1"))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)

View(trait_triples$joined_triples)

df<- trait_triples$joined_triples %>% 
      mutate(term1 = ifelse(term1 == 'Fatty Acid-Binding Protein 1', "FABP4", term1))
KEY_TERM= 'FABP4'


out <- overlap_trait_and_bc(df, KEY_TERM =KEY_TERM, n_filter=1, sankey_font = 13,bc_triples) # NB using n=1
out$full_sankey
out$subset_sankey

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```

## FGF7

```{r}

load(paste0(data_location, "lit_spaces_finalset_tidy.RData"))
id = 'prot-a-1097'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c('FGF7', "Fibroblast Growth Factor 7"))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)

View(trait_triples$joined_triples)

KEY_TERM= 'FGF7'


out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=1, sankey_font = 13,bc_triples) # NB using n=1
out$full_sankey
out$subset_sankey

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```






## IGF

```{r}

load(paste0(data_location, "lit_spaces_finalset_tidyV3.RData"))
id = 'ukb-d-30770_irnt'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c('Insulin-Like Growth Factor I', 'IGF1'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)

KEY_TERM= 'IGF1'


out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=2, sankey_font = 13,bc_triples)
out$full_sankey
out$subset_sankey

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```

## cardiotrophin 1

```{r}
load(paste0("../results/literature_outputs/lit_spaces_finalset_tidyV3.RData"))
id = 'prot-a-710'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_triples <- extract_two_triples_for_trait(trait_tidy,   KEY_TERM = c('cardiotrophin 1'))#, ignore_terms = c("AKT1"))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'cardiotrophin 1'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,
                                 bc_triples)


write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```


## IL1RL1

```{r}

load(paste0(data_location, "lit_spaces_finalset_tidy.RData"))
id = 'prot-b-38'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c( 'IL1RL1'))



trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM= 'IL1RL1'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=2, sankey_font = 13,bc_triples)
out$full_sankey
out$subset_sankey

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```



## IL3


```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-1530',  KEY_TERM = c('Interleukin-3 Receptor', 'IL3RA', 'Interleukin-3'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)


KEY_TERM = 
  
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=1,sankey_font = 13,
                                 bc_triples)

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```


## Alcohol dehydrogenase

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-67',  KEY_TERM = c('Alcohol dehydrogenase', 'Alcohol dehydrogenase (acceptor)'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)


KEY_TERM = "Alcohol dehydrogenase"
  
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=1,sankey_font = 13,
                                 bc_triples)

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```




```{r}

```


## ICAM1

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-1397',  KEY_TERM = c('Intercellular adhesion molecule 1', 
                                                                               "ICAM1",
                                                                               'Intercellular cell adhesion molecule', 
                                                                               'Intercellular Adhesion Molecules'  ))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM= 'ICAM1'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM = KEY_TERM , n_filter=2,
                                 bc_triples, sankey_font=13)

```







## Ferritin

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-1148',  KEY_TERM = c('Ferritin'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'Ferritin'
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,
                                 bc_triples)

```


## CRP

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-670',  KEY_TERM = c('CRP',  'C-reactive protein'))#, keep_only = c("CD40", "CD40 Ligand"))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>8), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'C-reactive protein'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM ='C-reactive protein' , n_filter=1,
                                 bc_triples)

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```

## calpastatin

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-366',  KEY_TERM = c('calpastatin'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'calpastatin'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,sankey_font = 13,
                                 bc_triples)

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```

## FGF7

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-1097',  KEY_TERM = c('FGF7', 'Fibroblast Growth Factor 7'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>2), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'FGF7'


out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,
                                 bc_triples)


write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```


## IL17RA


```{r}

trait_triples <- extract_two_triples_for_trait(id="prot-a-1486",  KEY_TERM = c('IL17RA'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 2), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'IL17RA'
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,
                                 bc_triples)

```


## Albumin

```{r}

trait_triples <- extract_two_triples_for_trait(id="met-c-841",  KEY_TERM = c('ALB', "Albumins"))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>10), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 50), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'Albumin'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM = KEY_TERM , n_filter=5,
                                 bc_triples, sankey_font=13)

```

## ANGPTL4

```{r}

trait_triples <- extract_two_triples_for_trait(id="prot-a-99",  KEY_TERM = c('ANGPTL4'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'ANGPTL4'
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM = KEY_TERM , n_filter=1,
                                 bc_triples)

```

```{r}

trait_triples <- extract_two_triples_for_trait(id="prot-a-1530",  KEY_TERM = c('Interleukin-3 Receptor', 'IL3RA'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'IL3RA'
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM = KEY_TERM , n_filter=1,
                                 bc_triples)

```


# Lifestyle traits

For lifestyle traits, the process is similar.

 - we load tidy trait space, by specifying whether it is a single-ID trait, or a combined trait: data is loaded from the selected RData
 - for lifestyle trait there is no anchor, so we first only extract a single triple from the trait (it is allowes to loop back on itself)
 - we also make a trait-only sankey 
 - then we connect BC space to trait space via that one triple A-B  and allow trait triple to backwards connect to another triple X-A
 - lit overlap sankey
 - save intermediates



```{r}
get_lifestyle_trait <- function(space, trait_id=NA, name=NA ) {
  data_location = '../results/literature_outputs/'
  
   if (space == 'single' & !is.na(trait_id)){
      load(paste0(data_location,"lit_spaces_finalset_tidyV3.RData"))
      trait_tidy <- tidy_litspace[[trait_id]] 
      rm(tidy_litspace)
      
    } else if (space == "combined" & !is.na(name)){
      load(paste0(data_location,"lit_spaces_combined_traits_tidyV3.RData"))
      trait_tidy <- tidy_combined_litspace[[name]]
      rm(tidy_combined_litspace)
    } 
  return(trait_tidy)
}
  
```

## HDL combined

```{r}
# not working
space = 'combined'
name='hdl_particles'

trait_tidy <- get_lifestyle_trait(space, trait_id=NA, name )

trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c('High Density Lipoproteins'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)

KEY_TERM= 'High Density Lipoproteins'


out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=2, sankey_font = 13,bc_triples)
out$full_sankey
out$subset_sankey

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))




####
trait_triples <- extract_lifestyle_main_triples(trait_tidy)


trait_san<- make_sankey(trait_triples %>% filter(n>3), fontSize=13)


overlap <- overlap_lifestyle_trait_and_bc(trait_triples, bc_triples,  n_filter=2)

sankey = make_sankey(overlap$sankey_data, colour_links = T, fontSize = 1, trait_col = "#7AC175", shared="#EF78E6")
sankey_sub = make_sankey(overlap$sankey_data_filtered, colour_links = T, fontSize = 13, trait_col = "#7AC175", shared="#EF78E6")


KEY_TERM="menopause"
write_tsv(as_tibble(overlap$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```


## childhood obesity

```{r}
space = 'single'
trait_id='ieu-a-1096'
trait_tidy <- get_lifestyle_trait(space, trait_id )


trait_triples <- extract_lifestyle_main_triples(trait_tidy)

trait_san<- make_sankey(trait_triples %>% filter(n>3), fontSize=13)


overlap <- overlap_lifestyle_trait_and_bc(trait_triples, bc_triples)
sankey = make_sankey(overlap$sankey_data, colour_links = T, fontSize = 12)


KEY_TERM="childhood_obesity"
write_tsv(as_tibble(overlap$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), "_NEW.csv"))


```


## sleep duration

```{r}

space = 'single'
trait_id='ukb-b-4424'
trait_tidy <- get_lifestyle_trait(space, trait_id )

trait_triples <- extract_lifestyle_main_triples(trait_tidy)

trait_san<- make_sankey(trait_triples %>% filter(n>0), fontSize=13)


overlap <- overlap_lifestyle_trait_and_bc(trait_triples, bc_triples)
sankey = make_sankey(overlap$sankey_data, colour_links = T, fontSize = 13)


KEY_TERM="sleep_duration"
write_tsv(as_tibble(overlap$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```


## snoring
```{r}
space = 'single'
trait_id="ukb-a-14"

apneas_variations <- c("Sleep Apnea Syndromes", "Sleep Apnea, Obstructive", "Sleep Apnea, Central", "OSA")

trait_tidy <- get_lifestyle_trait(space, trait_id ) 

trait_tidy_mod <- trait_tidy %>% 
   mutate(term1 = ifelse(term1 %in% apneas_variations, "Sleep Apnea", term1)) %>% 
   mutate(term2 = ifelse(term2 %in% apneas_variations, "Sleep Apnea", term2)) 

trait_tidy %>% write_csv("~/Desktop/snoring_literature_space.csv")
  
# trait_triples <- extract_lifestyle_main_triples(trait_tidy) remove too much for this trait; going to do some steps manually


trait_triple1 <- trait_tidy %>% 
   select(1:5) %>% 
   mutate(term1 = ifelse(term1 %in% apneas_variations, "Sleep Apnea", term1)) %>% 
   mutate(term2 = ifelse(term2 %in% apneas_variations, "Sleep Apnea", term2)) %>% 
    group_by(term1, predicate, term2) %>% 
    summarise_all(sum) %>% ungroup()
  
  print(trait_triple1 %>% select(term1, term2) %>% distinct() %>% dim())
  
trait_triples2 <- trait_triple1 %>% 
    tidy_terms_for_viz() %>% 
    select(term1,term2,n_pair) %>% 
    filter(term1 !=term2) %>% 
    group_by(term1, term2) %>% 
    slice(which.max(n_pair)) %>%
    ungroup()  

 # need to drop reverse connections: keep A-B or B-A depending which one is more common
  trait_triple2_onedir<- trait_triples2 %>% 
    # forward count
    rename(n_pair_f = n_pair) %>% 
    # join col that will show reverse pair count
    left_join(trait_triples2 %>% select(term1,term2,n_pair_b = n_pair), by = c('term1' = 'term2', 'term2' = 'term1')) %>% 
    distinct() %>% 
    # if rel does not exist uin reverse, set it to 0
    mutate(across(n_pair_b, ~replace_na(.x, 0))) %>% 
    # if f more common, keep it, else, keep reverse
    mutate(keep = ifelse(n_pair_f >= n_pair_b ,T,F)) %>% 
    filter(keep==T) %>%  rename(n_pair=n_pair_f) %>% 
    rename(n=n_pair)


trait_san<- make_sankey(trait_triple2_onedir %>% filter(n>0), fontSize=13)

htmlwidgets::saveWidget(trait_san, file=paste0( getwd(), "/snoring_literature_space.html"))


### overlap not working!
overlap <- overlap_lifestyle_trait_and_bc(trait_triple2_onedir, bc_triples)
sankey = make_sankey(overlap$sankey_data, colour_links = T, fontSize = 13)


# no
out <- overlap_trait_and_bc(trait_triple2_onedir, KEY_TERM = 'Sleep Apnea', n_filter=1, sankey_font = 13,bc_triples)


# mainstrem way
trait_triples <- extract_lifestyle_main_triples(trait_tidy_mod)

overlap <- overlap_lifestyle_trait_and_bc(trait_triples, bc_triples)
sankey = make_sankey(overlap$sankey_data, colour_links = T, fontSize = 12)




```




## menopause

```{r}
space = 'combined'
name='menopause'

trait_tidy <- get_lifestyle_trait(space, trait_id=NA, name )

trait_triples <- extract_lifestyle_main_triples(trait_tidy)


trait_san<- make_sankey(trait_triples %>% filter(n>3), fontSize=13)


overlap <- overlap_lifestyle_trait_and_bc(trait_triples, bc_triples,  n_filter=2)

sankey = make_sankey(overlap$sankey_data, colour_links = T, fontSize = 1, trait_col = "#7AC175", shared="#EF78E6")
sankey_sub = make_sankey(overlap$sankey_data_filtered, colour_links = T, fontSize = 13, trait_col = "#7AC175", shared="#EF78E6")


KEY_TERM="menopause"
write_tsv(as_tibble(overlap$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```







# testing path extraction 
```{r}

INTERMEDIATE= 'testosterone'

INTERMEDIATE= 'C-reactive protein'

INTERMEDIATE= 'SHBG'


X <- a_sub %>% filter(term1 == INTERMEDIATE | term2 == INTERMEDIATE)

Y <- X %>% filter(!(term1 == KEY_TERM | term2 == 'Breast Diseases' ))
y <- c(Y$term2)
y <- y[y !='Breast Diseases'  ]


Y1 <- a_sub %>% filter(term1 %in% Y$term2) %>%  filter(!term2 %in% y)
y1 <- c(Y1$term2, Y$term2)
y1 <- y1[y1 !='Breast Diseases'  ]

Y2 <- a_sub %>% filter(term1 %in% Y1$term2) %>%  filter(!term2 %in% y1)
y2 <- c(Y2$term2, Y1$term2, Y$term2)
y2 <- y2[y2 !='Breast Diseases'  ]

Y3 <- a_sub %>% filter(term1 %in% Y2$term2) %>%  filter(!term2 %in% y2)

test <- bind_rows(X, Y1, Y2, Y3)
test_sankey<- make_sankey(test, fontSize=18)

```
























