---
title: "R Notebook"
output: html_notebook
---

```{r message=F}
library(readr)
library(tidyr)
library(stringr)
library(vroom)
library(dplyr)
library(networkD3)

source("app3_sankey_app/functions_literature.R")

data_location = '../results/literature_outputs/'
```


# Breast cancer
```{r message=F}
# tidy breast cancer space and extract sequential triples

## prod data
dat<- vroom("../results/literature_outputs/breast_cancer_litspace_prod.csv",show_col_types = FALSE) %>%  mutate(lit.id = as.character(lit.id))

# tidy space 
bc_triples_tidy_count <- tidy_lit_space(dat) %>%    
    mutate(term1.type_verbose  = ifelse(grepl("obesity", term1, ignore.case = T), 'antro', term1.type_verbose)) %>%
    mutate(term2.type_verbose  = ifelse(grepl("obesity", term2, ignore.case = T), 'antro', term2.type_verbose))

write_tsv(bc_triples_tidy_count, "../results/literature_outputs/breast_cancer_litspace_prod_tidy_space.csv")

bc_triples <- get_breast_cancer_triples(bc_triples_tidy_count)
```



# Molecular traits 

For each molecular trait (i.e. trait with anchor):
	 - load tidy lit space form `lit_spaces_finalset_tidy.RData`
	 - extract and link trait triples
	 - make mini sankey of trait triples (for exploratory purposes)
	 - overlap literature spaces of trait and breast cancer
	 - the output has the full_sankey or subset_sankey (excludes n=1 (or more)) 
	 - save overlap intermediate terms


## IGF

```{r}

load(paste0(data_location, "lit_spaces_finalset_tidy.RData"))
id = 'ukb-d-30770_irnt'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_triples <- extract_two_triples_for_trait(trait_tidy,  KEY_TERM =c('Insulin-Like Growth Factor I', 'IGF1'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)

KEY_TERM= 'IGF1'


out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=2, sankey_font = 13,bc_triples)
out$full_sankey
out$subset_sankey

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```

## cardiotrophin 1

```{r}
load(paste0("../results/literature_outputs/lit_spaces_finalset_tidy.RData"))
id = 'prot-a-710'
trait_tidy <- tidy_litspace[[id]]
rm(tidy_litspace)


trait_triples <- extract_two_triples_for_trait(trait_tidy,   KEY_TERM = c('cardiotrophin 1'))#, ignore_terms = c("AKT1"))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'cardiotrophin 1'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,
                                 bc_triples)


write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```


## IL1RL1

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-b-38',  KEY_TERM = c('IL1RL1'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM= 'IL1RL1'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=1,sankey_font = 13,
                                 bc_triples)

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```



## IL3


```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-1530',  KEY_TERM = c('Interleukin-3 Receptor', 'IL3RA', 'Interleukin-3'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)


KEY_TERM = 
  
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=1,sankey_font = 13,
                                 bc_triples)

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```


## Alcohol dehydrogenase

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-67',  KEY_TERM = c('Alcohol dehydrogenase', 'Alcohol dehydrogenase (acceptor)'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)


KEY_TERM = "Alcohol dehydrogenase"
  
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM, n_filter=1,sankey_font = 13,
                                 bc_triples)

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```




```{r}

```


## ICAM1

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-1397',  KEY_TERM = c('Intercellular adhesion molecule 1', 
                                                                               "ICAM1",
                                                                               'Intercellular cell adhesion molecule', 
                                                                               'Intercellular Adhesion Molecules'  ))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM= 'ICAM1'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM = KEY_TERM , n_filter=2,
                                 bc_triples, sankey_font=13)

```







## Ferritin

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-1148',  KEY_TERM = c('Ferritin'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'Ferritin'
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,
                                 bc_triples)

```


## CRP

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-670',  KEY_TERM = c('CRP',  'C-reactive protein'))#, keep_only = c("CD40", "CD40 Ligand"))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>8), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'C-reactive protein'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM ='C-reactive protein' , n_filter=1,
                                 bc_triples)

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))

```

## calpastatin

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-366',  KEY_TERM = c('calpastatin'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'calpastatin'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,sankey_font = 13,
                                 bc_triples)

write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```

## FGF7

```{r}

trait_triples <- extract_two_triples_for_trait(id='prot-a-1097',  KEY_TERM = c('FGF7', 'Fibroblast Growth Factor 7'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>2), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'FGF7'


out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,
                                 bc_triples)


write_tsv(as_tibble(out$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```


## IL17RA


```{r}

trait_triples <- extract_two_triples_for_trait(id="prot-a-1486",  KEY_TERM = c('IL17RA'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 2), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'IL17RA'
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM =KEY_TERM , n_filter=1,
                                 bc_triples)

```


## Albumin

```{r}

trait_triples <- extract_two_triples_for_trait(id="met-c-841",  KEY_TERM = c('ALB', "Albumins"))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>10), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 50), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'Albumin'

out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM = KEY_TERM , n_filter=5,
                                 bc_triples, sankey_font=13)

```

## ANGPTL4

```{r}

trait_triples <- extract_two_triples_for_trait(id="prot-a-99",  KEY_TERM = c('ANGPTL4'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'ANGPTL4'
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM = KEY_TERM , n_filter=1,
                                 bc_triples)

```

```{r}

trait_triples <- extract_two_triples_for_trait(id="prot-a-1530",  KEY_TERM = c('Interleukin-3 Receptor', 'IL3RA'))

trait_trip_1s<- make_sankey(trait_triples$triple1_tidy %>% filter(n>0), fontSize=13)
trait_trip_2s<- make_sankey(trait_triples$triple2_tidy %>% filter(n > 0), fontSize=13)
trait_s_n<- make_sankey(trait_triples$joined_triples, fontSize=13)



KEY_TERM = 'IL3RA'
out <- overlap_trait_and_bc(trait_triples$joined_triples, KEY_TERM = KEY_TERM , n_filter=1,
                                 bc_triples)

```


# Lifestyle traits

For lifestyle traits, the process is similar.

 - we load tidy trait space, by specifying whether it is a single-ID trait, or a combined trait: data is loaded from the selected RData
 - for lifestyle trait there is no anchor, so we first only extract a single triple from the trait (it is allowes to loop back on itself)
 - we also make a trait-only sankey 
 - then we connect BC space to trait space via that one triple A-B  and allow trait triple to backwards connect to another triple X-A
 - lit overlap sankey
 - save intermediates



```{r}
get_lifestyle_trait <- function(space, trait_id=NA, name=NA ) {
  data_location = '../results/literature_outputs/'
  
   if (space == 'single' & !is.na(trait_id)){
      load(paste0(data_location,"lit_spaces_finalset_tidy.RData"))
      trait_tidy <- tidy_litspace[[trait_id]] 
      rm(tidy_litspace)
      
    } else if (space == "combined" & !is.na(name)){
      load(paste0(data_location,"lit_spaces_combined_traits_tidy.RData"))
      trait_tidy <- tidy_combined_litspace[[name]]
      rm(tidy_combined_litspace)
    } 
  return(trait_tidy)
}
  
```

## childhood obesity

```{r}
space = 'single'
trait_id='ieu-a-1096'
trait_tidy <- get_lifestyle_trait(space, trait_id )


trait_triples <- extract_lifestyle_main_triples(trait_tidy)

trait_san<- make_sankey(trait_triples %>% filter(n>3), fontSize=13)


overlap <- overlap_lifestyle_trait_and_bc(trait_triples, bc_triples)
sankey = make_sankey(overlap$sankey_data, colour_links = T, fontSize = 12)


KEY_TERM="childhood_obesity"
write_tsv(as_tibble(overlap$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```


## sleep duration

```{r}

space = 'single'
trait_id='ukb-b-4424'
trait_tidy <- get_lifestyle_trait(space, trait_id )

trait_triples <- extract_lifestyle_main_triples(trait_tidy)

trait_san<- make_sankey(trait_triples %>% filter(n>0), fontSize=13)


overlap <- overlap_lifestyle_trait_and_bc(trait_triples, bc_triples)
sankey = make_sankey(overlap$sankey_data, colour_links = T, fontSize = 13)


KEY_TERM="sleep_duration"
write_tsv(as_tibble(overlap$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```



## menopause

```{r}
space = 'combined'
name='menopause'

trait_tidy <- get_lifestyle_trait(space, trait_id=NA, name )

trait_triples <- extract_lifestyle_main_triples(trait_tidy)


trait_san<- make_sankey(trait_triples %>% filter(n>3), fontSize=13)


overlap <- overlap_lifestyle_trait_and_bc(trait_triples, bc_triples,  n_filter=2)

sankey = make_sankey(overlap$sankey_data, colour_links = T, fontSize = 1, trait_col = "#7AC175", shared="#EF78E6")
sankey_sub = make_sankey(overlap$sankey_data_filtered, colour_links = T, fontSize = 13, trait_col = "#7AC175", shared="#EF78E6")


KEY_TERM="menopause"
write_tsv(as_tibble(overlap$terms_list), paste0("../results/literature_outputs/sankey_terms_storage/lit_terms_", c(gsub( " ", "", KEY_TERM)), ".csv"))


```







# testing path extraction 
```{r}

INTERMEDIATE= 'testosterone'

INTERMEDIATE= 'C-reactive protein'

INTERMEDIATE= 'SHBG'


X <- a_sub %>% filter(term1 == INTERMEDIATE | term2 == INTERMEDIATE)

Y <- X %>% filter(!(term1 == KEY_TERM | term2 == 'Breast Diseases' ))
y <- c(Y$term2)
y <- y[y !='Breast Diseases'  ]


Y1 <- a_sub %>% filter(term1 %in% Y$term2) %>%  filter(!term2 %in% y)
y1 <- c(Y1$term2, Y$term2)
y1 <- y1[y1 !='Breast Diseases'  ]

Y2 <- a_sub %>% filter(term1 %in% Y1$term2) %>%  filter(!term2 %in% y1)
y2 <- c(Y2$term2, Y1$term2, Y$term2)
y2 <- y2[y2 !='Breast Diseases'  ]

Y3 <- a_sub %>% filter(term1 %in% Y2$term2) %>%  filter(!term2 %in% y2)

test <- bind_rows(X, Y1, Y2, Y3)
test_sankey<- make_sankey(test, fontSize=18)

```
























